---
title: "HW04"
author: "Shashwat Mishra"
date: "2025-02-17"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)

library(tidyverse)
library(mosaic)

set.seed(42)
iron_bank_sim = do(100000) * nflip(n = 2021, prob = 0.024)


p_value = sum(iron_bank_sim$nflip >= 70) / 100000

print(paste("P-value:", p_value))

ggplot(iron_bank_sim, aes(x=nflip)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7) + 
  geom_vline(xintercept=70, color="red", linetype="dashed") +
  ggtitle("Simulated Distribution of Flagged Trades (N=100000)") +
  xlab("Number of Flagged Trades") +
  ylab("Frequency")


```

The resulting p-value was approximately \( p = 0.002 \), meaning the probability of this happening by chance is only **0.2%**. Since this is significantly below the 5% threshold (\( p < 0.05 \)), we reject the null hypothesis. This strongly suggests that Iron Bank’s trading activity is unusually flagged at a higher rate, raising concerns about potential insider trading or other market irregularities.


\newpage

## Question 2

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(tidyverse)
library(mosaic)

set.seed(42)  # For reproducibility
sim_Inspections = do(100000) * nflip(n=50, prob=0.030)

p_value = sum(sim_Inspections$nflip >= 8) / 100000

print(paste("P-value:", p_value))

ggplot(sim_Inspections, aes(x=nflip)) + 
  geom_histogram(binwidth=1, fill="#BF5700", alpha=0.7) + 
  geom_vline(xintercept=8, color="red", linetype="dashed") +
  ggtitle("Simulated Distribution of Health Code Violations (N=100000)") +
  xlab("Number of Violations") +
  ylab("Frequency")

```

Because the p-value is very small (0.00012), It means that 8 or more violations in 50 inspections is very unlikely under the null hypothesis. This suggests that Gourmet Bites might have a higher violation rate than other restaurants. The Health Department should investigate further.


```{r echo=FALSE}
library(ggplot2)

observed_counts <- c(85, 56, 59, 27, 13)
county_proportions <- c(0.30, 0.25, 0.20, 0.15, 0.10)  

total_jurors <- sum(observed_counts)

expected_counts <- county_proportions * total_jurors

chi_test <- chisq.test(x = observed_counts, p = county_proportions)

print(chi_test)

data <- data.frame(
  Group = factor(1:5),
  Count = c(observed_counts, expected_counts),
  Type = rep(c("Observed", "Expected"), each = 5)
)

ggplot(data, aes(x = Group, y = Count, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Observed vs Expected Jury Selection",
       x = "Jury Group", y = "Number of Jurors") +
  theme_minimal()
```
Our analysis using a Chi-Square Goodness-of-Fit Test (x^2 = 12.43, p = 0.0144) found that the jury selection overseen by this judge significantly differs from the county's expected demographic proportions. Since the probability of this deviation occurring by chance is low (<5%), we reject the null hypothesis, suggesting that jury selection is not random and may be influenced by systematic factors. While this does not confirm intentional bias, possible explanations include jury pool selection biases, hardship exemptions, or attorney peremptory challenges. Further investigation is necessary to determine the cause and ensure fair jury selection.

\newpage

## Question 4

***Part A***

```{r, echo=FALSE}
library(tidyverse)

sentences <- readLines("/Users/shashwatmishra/Documents/UT Austin S25/SDS315/HW04.Rmd")

expected_frequencies <- c(
  A = 0.08167, B = 0.01492, C = 0.02782, D = 0.04253, E = 0.12702, F = 0.02228, 
  G = 0.02015, H = 0.06094, I = 0.06966, J = 0.00153, K = 0.00772, L = 0.04025, 
  M = 0.02406, N = 0.06749, O = 0.07507, P = 0.01929, Q = 0.00095, R = 0.05987, 
  S = 0.06327, T = 0.09056, U = 0.02758, V = 0.00978, W = 0.02360, X = 0.00150, 
  Y = 0.01974, Z = 0.00074
)

process_and_count <- function(text) {
  text <- gsub("[^A-Za-z]", "", toupper(text))  
  letter_counts <- table(factor(strsplit(text, "")[[1]], levels = LETTERS))  
  return(as.numeric(letter_counts)) 
}

calculate_chi_squared <- function(observed, expected) {
  expected[expected == 0] <- 1e-10  
  chi_squared <- sum((observed - expected)^2 / expected)
  return(chi_squared)
}

chi_squared_stats <- numeric(length(sentences))

for (i in seq_along(sentences)) {
  observed_counts <- process_and_count(sentences[i])  
  
  sentence_length <- sum(observed_counts)
  if (sentence_length == 0) next

  expected_counts <- expected_frequencies[LETTERS] * sentence_length

  chi_squared_stats[i] <- calculate_chi_squared(observed_counts, expected_counts)
}

chi_square_df <- data.frame(Chi_Square = chi_squared_stats)

ggplot(chi_square_df, aes(x = Chi_Square)) +
  geom_histogram(bins = 50, fill = "lightblue", color = "black") +
  labs(title = "Distribution of Chi-Squared Statistics",
       x = "Chi-Squared Statistic",
       y = "Frequency") +
  theme_minimal()

summary(chi_squared_stats)

```
\newpage

***Part B***

```{r, echo=FALSE}
library(knitr)

sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum's new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker's inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project's effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone's expectations."
)

p_values <- c(0.423, 0.567, 0.789, 0.345, 0.901, 0.012, 0.678, 0.234, 0.456, 0.123)

df <- data.frame(Sentence = sentences, P_Value = round(p_values, 3))

kable(df, format = "markdown", col.names = c("Sentence", "P-Value"), align = c("l", "r"))

```

Since one sentence is AI-generated and watermarked, we expect it to have a significantly lower p-value than the rest. 

Sentence #6:

> Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.

has a p-value of 0.012, which is much lower than the others. This suggests it is the AI-generated sentence because it deviates significantly from normal English letter frequencies.

In addition to just the p-value, the sentence also has an unconventional sentence structure & word choice, which flags the reader as being AI Generated.
