---
title: "HW08"
author: "Shashwat Mishra | sdm4329"
date: https://github.com/shashwat-m/SDS315
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(mosaic)
library(dplyr)
library(ggplot2)
```


\newpage

# Question 1


## Part A
```{r message=FALSE, echo=FALSE, warning=FALSE, output = FALSE, results='hide'}
creatinine <- read.csv("/Users/shashwatmishra/Downloads/creatinine.csv")
max_count <- max(creatinine$creatclear)
min_count <- min(creatinine$creatclear)
mean <- mean(creatinine$creatclear)

lin_model <- lm(creatclear ~ age, data=creatinine)
summary(lin_model)

creat_55 <- 147.81292 + (-0.61892 * 55)
```

The expected creatinine clearance rate for someone who is 55 years old is `r round(creat_55, 3)`. To find this, I used a linear model using the lm() function to find the coefficients of the slope and intercept of the line of best fit of the data. The coefficients I got were -0.61982 for the age, and 147.81292 for the intercept. Using this, I made the formula -0.61982(age) + 147.81292 to find the expected creatinine level of a specific age. 


## Part B

The creatinine clearance rate decreases by approximately 0.62 mL/minute per year. I determined this by examining the slope of the linear regression line, which represents the change in clearance rate per one-year increase in age.


## Part C

```{r echo=FALSE}
creat_40 <- 147.81292 + (-0.61892 * 40)
creat_60 <- 147.81292 + (-0.61892 * 60)
```

The 40 year old with a clearance of 135 mL/minute is healthier than the 60 year old with a clearance of 112 ml/minute because according to the regression model, the expected for both the 40 and 60 year old is 123.05 mL/minute and 110.68 mL/minute, respectively. Based on this, we can see that the 40 year old's expected is far larger than the expected, overtaking it by about 12 mL/minute, whereas the 60 year old is over by only about 2 mL/minute. 

Because we know that higher clearances are better, it's safe to say that the 40 year old's clearance is considerably healthier.

# Question 2

```{r pressure, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
covid <- read.csv("/Users/shashwatmishra/Downloads/covid.csv")

italy <- covid[covid$country == "Italy",]
spain <- covid[covid$country == "Spain",]


model_italy <- lm(log(deaths) ~ days_since_first_death, data = italy)
model_spain <- lm(log(deaths) ~ days_since_first_death, data = spain)

summary(model_italy)

k_italy <- coef(model_italy)[2]
k_spain <- coef(model_spain)[2]

doubling_italy <- log(2) / k_italy
doubling_spain <- log(2) / k_spain

boot_italy <- do(10000) * {
  m <- lm(log(deaths) ~ days_since_first_death, data = resample(italy))
  k <- coef(m)[2]
  data.frame(growth = k, doubling = log(2) / k)
}

# Spain bootstrap
boot_spain <- do(10000) * {
  m <- lm(log(deaths) ~ days_since_first_death, data = resample(spain))
  k <- coef(m)[2]
  data.frame(growth = k, doubling = log(2) / k)
}


summary(boot_italy$growth)
summary(boot_italy$doubling)

# Spain CI (approximate)
summary(boot_spain$growth)
summary(boot_spain$doubling)


ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(
    title = "Daily COVID-19 Deaths Over Time in Italy and Spain",
    x = "Days Since First Reported Death",
    y = "Daily Reported Deaths",
    color = "Country"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

The estimated growth rate and doubling time for Italy were 0.183 and 3.78 days, respectively.
For Spain, the estimated growth rate was 0.276, with a corresponding doubling time of 2.51 days

# Question 3


```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

milk <- read.csv("/Users/shashwatmishra/Downloads/milk.csv")
milk_model <- lm(sales ~ price, data=milk)

summary(milk_model)
price_slope <- coef(milk_model)[2]

log_model <- lm(log(sales) ~ log(price), data = milk)
summary(log_model)
elasticity <- coef(log_model)[2]

library(mosaic)

set.seed(123)
boot_elasticity <- do(10000) * {
  m <- lm(log(sales) ~ log(price), data = resample(milk))
  data.frame(beta = coef(m)[2])
}

quantile(boot_elasticity$beta, c(0.025, 0.975))

        
```

To estimate the price elasticity of demand for milk, I fit a log-log linear model of log(sales) on log(price), consistent with a power-law demand model.

The estimated price elasticity was approximately -1.61, with a 95% bootstrapped confidence interval from -1.77 to -1.46.

This suggests that a 1% increase in the price of milk leads to a 1.46% to 1.77% decrease in the quantity of milk purchased, indicating elastic demand.



